\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-07-11 ResponseToReview.tex

10) Materials and Methods:
Page 5, lines 22 - 23: how many individuals per accession? A fixed number, or variable, and depending on what? This can be mentioned so the reader does not have to go to the appendix or supplements.\\

\res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}

11) Page 5, Line 28: mention how big the hopscotch element is.  Can it be amplified in one amplicon? Again, this would be useful to know without having to look in appendix or supplements.\\

\res{The size of the amplicon is listed in the Materials and Methods following the primer sequences: ``Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."}

@@

19) Page 9 lines 9 - 10: "We genotyped individuals for the Hopscotch insertion following the protocols listed above." The PCR protocols or the sequencing protocols?\\

\res{We have clarified this in the manuscript.}

20) Results:
Page 10 lines 5 - 6: "Within our parviglumis and mexicana samples we found the Hopscotch insertion segregating in 37 and four populations, respectively"; remind us how many parviglumis and Mexicana samples you had total? So we can see if this comes out to a very different ratio than in maize.\\

\res{We have added ratios to the text: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (37 out of 86) and four (4 out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico.}"

@@

21) Page 10 lines 7 - 9: "Using our Hopscotch genotyping, we calculated diﬀerentiation between populations (FST) and subspecies (FCT) for populations in which we sampled eight or more alleles." What do you mean, 8 or more alleles? I thought you only had presence or absence; how are alleles differentiated with your PCR?\\

\res{We believe we addressed this with editing for your previous comment concerning uncertainty of the PCRs we were using. As edited into the Materials and Methods section, there is a PCR that produces a specific size for each allele. We have edited subsequent text accordingly.}

### 2014-07-11 manuscript.tex
Edits
working on reviewer 2 still. Edited .bib file — need to figure out how
to make it recompile.
edits from reviewer 2
edits to sampling details

\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article). DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}. We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion. Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR). Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'}; 
HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article). Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $\sim$1.1kb), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article). Since we had a PCR for each allele, if only one PCR resolved well, we scored one allele for the individual. We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min. PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.

@@

Starting on day 15, we measured tillering index as the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}. Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60. On day 65 we measured culm diameter between the third and fourth nodes of each plant. Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol. We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above. Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from \citet{Briggs2007}, which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from \citep{Briggs2007}.

@@

\textbf{Genotyping---}Genotype of the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals. Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article). Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and four populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig. \ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals. We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}). Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).

>>>>

\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article). DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}. We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion. Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR). Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'}; 
HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article). Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article). Since we had a PCR for each allele, if only one PCR resolved well, we scored one allele for the individual. We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min. PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.

@@

Starting on day 15, we measured tillering index as the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}. Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60. On day 65 we measured culm diameter between the third and fourth nodes of each plant. Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol. We genotyped individuals for the \emph{Hopscotch} insertion following the PCR protocols listed above. Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from \citet{Briggs2007}, which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from \citep{Briggs2007}.

@@

\textbf{Genotyping---}Genotype of the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals. Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article). Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (37 out of 86) and four (4 out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig. \ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals. We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}). Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).
